Graph Testing

### 2020-03-12 notes/graph_testing.tex
updated summary documents

@@

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

@@

We establish some notation that we will use throughout. A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $ij$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$. We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with. When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.

@@

We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
\end{equation*}
The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
\end{equation*}

@@

\paragraph{Eigenvector projection test statistic.}
The graph Laplacian eigenvector projection test is a truncated-series test. Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.  Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic. Formally,
\begin{equation*}
T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
\end{equation*}
Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.

@@

\begin{theorem}
	\label{thm:sobolev_testing_rate_order1}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ is bounded away from zero and infinity, 
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
	\begin{equation*}
	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$. Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); H^1(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$. In the case where $f \in H^1(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$. We shall discuss testing in the low-smoothness regime shortly. First, let's consider the situation when we assume more derivatives worth of smoothness.

@@

When we additionally believe $H^s(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$. Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $H^s(\mathcal{X})$. However, if we assume $f$ and its derivatives are compactly supported on $\Xset$, we recover the usual rate. Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $H_0^s(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.

\begin{theorem}
	\label{thm:sobolev_testing_rate}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e. there exists $p_{\min} > 0$ such that 
	\begin{equation*}
	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
	\begin{equation*}
	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H_0^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

@@

The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
\begin{equation*}
H^{s}(\mathcal{X}) \subseteq \Leb^4(\mathcal{X}) 
\end{equation*}
does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
\begin{proposition}
	\label{prop:L4_testing_rate}
	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
	\begin{equation}
	\label{eqn:L4_testing_rate}
	\epsilon^2 > b \cdot n^{-1/2}
	\end{equation}
	then the test
	\begin{equation*}
	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
	\end{equation*}
	has worst-case risk
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H^1(\mathcal{X};R)) \lesssim \frac{1}{b}. 
	\end{equation*}
\end{proposition}
Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.

@@

To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk. Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
\begin{equation}
y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
\end{equation}
for some fixed $\beta \in \Reals^n$.  In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$. Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.

>>>>

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

@@

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

@@

We establish some notation that we will use throughout. A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $(ij)$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$. We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with. When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.

@@

We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
\end{equation*}
(Not that the expectation here may taken over the random responses $Y$ when $X$ is assumed to be fixed, or jointly over the randomness of $(X,Y)$ when $X$ is assumed random.) The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
\end{equation*}
For a given level $\alpha$ of tolerated error, the minimax critical radius
\begin{equation*}
\epsilon^{\star}(\mathcal{H},\alpha) := \inf \Bigl\{\epsilon > 0: \inf_{\phi} R_{\epsilon}(\mathcal{H}) \leq \alpha \Bigr\}
\end{equation*}
then provides a lens through which one may examine the hardness of testing over $\mathcal{H}$. We will sometimes suppress the dependence of the minimax critical radius on $\alpha$ and simply write $\epsilon^{\star}(\mathcal{H})$.

@@

\paragraph{Eigenvector projection test statistic.}
The graph Laplacian eigenvector projection test is a truncated-series test. Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.  Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic. Formally, let
\begin{equation*}
\Pi_{\kappa,G}(f) := \frac{1}{n}\sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} f(x_i) v_{k,i}(G)\Biggr) v_{k}(G),~~ T_{\mathrm{spec}}(G) := \norm{\Pi_{\kappa,G}(f)}_n^2
\end{equation*}
where $\norm{\theta}_n^2 := \frac{1}{n} \sum_{i = 1}^{n} \theta_i^2$ for $\theta \in \Reals^n$.
Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.

@@

\begin{theorem}
	\label{thm:sobolev_testing_rate_order1}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ bounded away from zero and infinity, 
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
	\begin{equation*}
	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$. Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); H^1(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$. In the case where $f \in H^1(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$. We shall discuss testing in the low-smoothness regime shortly. First, let's consider the situation when we are willing assume $f$ possesses additional derivatives bounded in $\Leb^2$ norm.

@@

When we are additionally willing to assume $H^s(\Xset)$, we know (again from \textcolor{red}{Ingster}) that the minimax testing rate is $n^{-4s/(4s + d)}$. Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $H^s(\mathcal{X})$. However, if we assume $f$ and its derivatives are compactly supported on $\Xset$, we recover the usual rate. Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $H_0^s(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.

\begin{theorem}
	\label{thm:sobolev_testing_rate}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $R > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e. there exists $p_{\min} > 0$ such that 
	\begin{equation*}
	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
	\begin{equation*}
	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = (Rn)^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H_0^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

@@

The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
\begin{equation*}
H^{s}(\mathcal{X}) \subseteq \Leb^4(\mathcal{X}) 
\end{equation*}
does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
\begin{proposition}
	\label{prop:L4_testing_rate}
	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
	\begin{equation}
	\label{eqn:L4_testing_rate}
	\epsilon^2 > c^2 b \cdot n^{-1/2}
	\end{equation}
	then the test
	\begin{equation*}
	\phi_{\mathrm{mean}} = \1\Bigl\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\Bigr\}
	\end{equation*}
	has worst-case risk
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H^1(\mathcal{X};R)) \lesssim \frac{1}{b}. 
	\end{equation*}
\end{proposition}
Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.

\subsection{Fixed Design}
Frequently in nonparametric regression problems, it is assumed that the design points are fixed, and, in the simplest case, uniformly spaced on the unit cube. For convenience in this setting we let $n = N^d$ for some integer $N$. The fixed grid design $\wb{X}$ consists of $n$ total evenly spaced grid points on $[0,1]^d$,
\begin{equation*}
\wb{X} := \Bigl\{\Bigl(\frac{2i_1 - 1}{2N},\ldots,\frac{2i_d - 1}{2N}\Bigr): i \in [N]^d\Bigr\}
\end{equation*}
Suppose we observe $n$ samples according to regression model
\begin{equation}
\label{eqn:grid_regression_model}
y_i = f(\wb{x}_i) + \varepsilon_i, ~\varepsilon_i \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $i \in [N]^d$,}
\end{equation}
Clearly model~\eqref{eqn:grid_regression_model} bears some similarity with model~\eqref{eqn:regression_random_design_known_variance}, the main difference being that the design $\wb{X}$ is now fixed instead of random.

We study the behavior of $\phi_{\spec}(\wb{G})$, where $\wb{G}$ is the lattice
\begin{equation*}
\wb{G} := \bigl(\wb{X},\wb{E}\bigr),~~ \wb{E} := \set{(\wb{x}_i,\wb{x}_j): i,j \in [N]^d, \norm{i - j}_1 = 1},
\end{equation*}
a very natural graph to build given our assumed grid design. Note that for a given $d$, we also have the recursive relation
\begin{equation*}
\wb{G} := \wb{G}_1 \otimes \ldots \otimes \wb{G}_1
\end{equation*}
where the tensor product is taken $d$ total times, and $\wb{G}_1$ is the $1$d lattice (i.e. the path) graph on $N$ vertices. For this reason it will be convenient to change notation slightly, and index by $d$-tuples of numbers rather than by numbers. In particular, we index the eigenvalues and eigenvectors of $L_{\wb{G}}$ as
\begin{equation*}
\lambda_k(\wb{G}) = \prod_{j = 1}^{d} \lambda_{k_j}(\wb{G}_1), v_{k,i}(\wb{G}) = \prod_{j = 1}^{d} v_{k_j,i_j}(\wb{G}_1)~~\textrm{for $k,i \in [N]^d$.}
\end{equation*}
Then, for a given $\kappa \in [n]$, letting $K = \kappa^{1/d}$, we see that our graph spectral test statistic may be written equivalently as
\begin{equation*}
T_{\spec}(\wb{G}) = \frac{1}{n} \sum_{k \in [K]^d} \Biggl(\sum_{i \in [N]^d} y_i v_{k,i}(G)\Biggr)^2
\end{equation*}

We see the test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Holder ball $C^{1}(\Xset;L)$ when $d \leq 4$, as the test $\phi_{\spec}(G_{n,r})$ does over the Holder ball $C^1(\Xset;L)$.
\begin{theorem}
	\label{thm:holder_testing_rate_grid}
	Let $L > 0$ be a fixed constant, and let $d \leq 4$. Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}. Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
	\begin{equation*}
	\kappa = (Ln)^{2d/(4 + d)},~~ \tau(b) = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for 
	\begin{equation}
	\label{eqn:holder_testing_rate_grid_1}
	\epsilon^2 \geq c \cdot \max\{L^{d/(4+d)}, L^{d/(4 + d) + 1}\} \cdot b n^{-4/(4 + d)}
	\end{equation}
	the worst-case risk of $\phi_{\spec}(\wb{G})$ on $C^{1}(\Xset;L)$ is upper bounded
	\begin{equation}
	\label{eqn:holder_testing_rate_grid_2}
	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C^{1}(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
	\end{equation}
\end{theorem}

As was the case with the neighborhood graph in the random design setting, the grid Sobolev norm fails to track the higher order derivatives of $f$ near the boundary of $\Xset$. We therefore will again insist that $f$ be compactly supported within $\Xset$. Under this condition, we find that the test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Holder ball $C_c^{s}(\Xset;L)$ when $4s \geq d$ as did the test $\phi_{\spec}(G_{n,r})$ over the Holder ball $C_c^s(\Xset;L)$.

\begin{theorem}
	\label{thm:holder_testing_rate_grid_higher_order}
	Let $L > 0$ be a fixed constant, $s \geq 2$ be a fixed integer, and let $d \leq 4s$.  Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}. Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
	\begin{equation*}
	\kappa = (Ln)^{2d/(4s + d)},~~ \tau(b) =  \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*} 
	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
	\begin{equation*}
	\epsilon^2 \geq c L^2 b n^{-4s/(4s + d)},
	\end{equation*}
	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
	\begin{equation*}
	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C_c^{s}(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
	\end{equation*}
\end{theorem}

Roughly speaking, we have shown that in the high-smoothness regime $4s > d$, testing over Holder classes in the grid design and random design cases is an equally difficult problem from the minimax perspective. The same is not true for the corresponding Sobolev spaces. When $2s > d$, the Sobolev spaces $H^s(\Xset;L)$ embed compactly into (lower-order) Holder classes, implying that every equivalence class $u \in H^s(\Xset;L)$ contains a favored representer which is bounded and Holder continuous, with bounded Holder norm. When $2s < d$, the situation is less nice; the spaces $H^s(\Xset;L)$ now contain (equivalence classes of) bump functions of arbitrarily large height and small width. By placing such bump functions at design points, it is possible to make the sampling model~\eqref{eqn:grid_regression_model} contain arbitrarily little information about the overall behavior of the function $f$ over $\Xset$. It is not hard to construct a function $f \in H^s(\Xset;L)$ for which, say, $\norm{f}_{\Leb^2(\Xset)} = 1$ yet $f(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$, implying that the critical radius $\epsilon^{\star}(H^s(\Xset;L)) = \Omega(1)$. We note that this is not an issue in the random design setting, because the design points are (randomly) chosen only after the function $f$ is fixed.

\textcolor{red}{(TODO): Make this discussion more rigorous. Perhaps appeal to the concept of precise representatives defined in (Evans and Gariepy).}

The above discussion leaves hope that when $2s > d$, consistent testing might still be possible. We can in fact go beyond this and show that the graph spectral test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Sobolev space $\wt{H}^s(\Xset;L)$ as the test $\phi_{\spec}(G_{n,r})$ did over $H_0^s(\Xset;L)$.

\begin{theorem}
	\label{thm:sobolev_testing_rate_grid}
	Let $L > 0$ be a fixed constant, $s \geq 1$ be a fixed integer, and let $2s > d$.  Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}. Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
	\begin{equation*}
	\kappa = n^{2d/(4s + d)},~~ \tau(b) =  \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*} 
	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
	\begin{equation*}
	\epsilon^2 \geq c L^2 b n^{-4s/(4s + d)},
	\end{equation*}
	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_grid_1}
	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});\wt{H}^s(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}
	\end{equation}
	for all $n$ sufficiently large.
\end{theorem}

\subsubsection{Low-smoothness regime with fixed design}
In the low-smoothness regime $d > 4s$, the coupling between empirical norm over $\wb{X}$ and the continuum $\Leb^2$ norm is sufficiently weak that the discretization error can become the dominant source of error, as we show in Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}.

\begin{lemma}
	\label{lem:holder_testing_rate_grid_low_smoothness_lb}
	For any $L > 0$ and any integers $s$ and $d$, there exists a function $g \in C^s(\Xset;L)$ such that
	\begin{equation*}
	\norm{g}_{\Leb^2}^2 \geq \frac{L^2}{2^d \pi^{2s} (d + 1)^{2s}}n^{-2s/d} 
	\end{equation*}
	but $g(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$. 
\end{lemma}
Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} implies the critical radius over the Holder classes $C^s(\Xset;L)$ must be at least on the order of $\epsilon^{\star}(C^s(\Xset;L)) \gtrsim n^{-s/d}$. When $4s \geq d$, this term is neglible relative to $n^{-2s/(4s + d)}$, and we arrive at the ``usual'' rates of minimax testing we see in Theorems~\ref{thm:holder_testing_rate_grid} and~\ref{thm:holder_testing_rate_grid_higher_order}. On the other hand when $4s < d$, this becomes the dominant term, and we can no longer achieve the typical rate $\epsilon^2 \asymp n^{-4s/(4s + d)}$.

In this setting, when we enforce the wider radius $\epsilon^2 \asymp n^{-2s/d}$, Proposition~\ref{prop:holder_testing_rate_grid_low_smoothness_ub} shows that a very simply test has small worst-case risk $\mathcal{R}_{\epsilon}$.

\begin{proposition}
	\label{prop:holder_testing_rate_grid_low_smoothness_ub}
	Let $L > 0$ be a fixed constant, $s \geq 1$ a fixed integer, and let $d \geq 4s$. Suppose we observe data according to the grid design regression model~\eqref{eqn:grid_regression_model}. Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
	\begin{equation*}
	\kappa = n,~~ \tau(b) = 1 + b\sqrt{\frac{2}{n}}
	\end{equation*}
	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any 
	\begin{equation*}
	\epsilon^2 \geq c b L^2 n^{-2s/d}
	\end{equation*}
	the worst case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\phi_{\spec}(\wb{G}); C^s(\Xset;L)) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2n}}
	\end{equation*}
\end{proposition}

\begin{itemize}
	\item The test statistic $T_{\spec}(\wb{G})$ is simply the empirical norm of $Y$. 
	\item Together, Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} and Proposition~\eqref{prop:holder_testing_rate_grid_low_smoothness_ub} characterize the minimax rate of the grid design regression testing problem when $4s > d$.
	\item This minimax rate does not match the upper bound established in Proposition~\ref{prop:L4_testing_rate}---the fixed design problem is harder. 
	\item Note that the usual minimax rate of estimation error over Holder classes is always $n^{-2s/(2s + d)}$, regardless of the relation between $s$ and $d$, and regardless of whether loss is measured in $\Leb^2$-norm or empirical norm. The explanation for this is that the rate $n^{-2s/(2s + d)}$ is always larger than $n^{-2s/d}$, and so the discretization error is never the dominant source of error in estimation. This reveals an interesting distinction between the testing and estimation problems, as testing is easy enough in a statistical sense that error incurred by the discrete nature of the problem may be the bottleneck.
\end{itemize}

@@

To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk. Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
\begin{equation}
\label{eqn:fixed_graph_regression_model}
y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
\end{equation}
for some fixed $\beta \in \Reals^n$.  In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$. Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
